Symmetric Peer-to-Peer Applications

### 2022-11-29 final/p2p-22.tex
- performance comparison

In this section, we revisit existing solutions for symmetric distributed
applications.
We focus on
    (a) how the network is organized,
    (b) how global time is determined, and
    (c) how events are propagated and applied.
Figure~\ref{fig.table} compares three selected works regarding these aspects.

@@

We now discuss related work on time machines.
%
Regarding decentralized solutions, Fusion~\cite{fusion} is a state
synchronization networking library for the Unity game engine.
It provides networked tick-based simulation in two modes: client-server
(\emph{hosted mode}) and peer-to-peer (\emph{shared mode}).
The hosted mode is based on continuous memory snapshots, which allows for
full rollbacks when clients diverge.
However, the shared mode is less powerful and only supports eventual
consistency among clients.
We presume that continuous memory snapshots would be too costly without a
central server.
%
Regarding single-user local machines, the video game Braid%
\footnote{Braid: \url{http://braid-game.com/}}
is designed on the assumption that players have unrestricted rollbacks as part
of the game mechanics.
Like Fusion, the implementation is based on continuous memory snapshots,
instead of deterministic simulation as we propose in this work.
%
Schuster~\cite{tml.js} proposes a simple API for time travelling in
event-driven JavaScript:
    \code{init()} to return an initial state;
    \code{handle(evt,mem)} to process events and modify state; and
    \code{render(mem)} to output the current state.
As we describe next, we use a similar approach by separating state mutation
from rendering~\cite{tml.alive}.

>>>>

In this section, we revisit existing solutions for symmetric distributed
applications.
We focus on
    (a) how the network is organized,
    (b) how global time is determined, and
    (c) how events are propagated and applied.
Figure~\ref{fig.table} compares three selected works regarding these aspects.
%
At the end of this section, we also discuss related work on local and
distributed time machines.

\subsection{Symmetric Distributed Applications}

@@

Many aspects may affect the performance of symmetric distributed applications,
such as
    the network latency,
    the number of nodes in the network,
    the rate of user events,
    and the application frame rate.
%
It is expected that centralized solutions exhibit a better performance in
comparison to peer-to-peer applications, since nodes are only 2 hops aways from
each other.
%
The performance of Croquet depends exclusively on the network latency, which
restricts the application to a maximum FPS.
For instance, since all frame ticks need to be transmitted over the network,
a latency of 50ms supports at most 20 FPS.
Also, the higher is the FPS, the higher is the network traffic and server load,
which may also degrade performance.
%
GALS detaches latency from FPS, since each node has its own timeline.
However, the protocol requires two roundtrips to the server, which degrades
performance considerably.
The authors claim to achieve a responsiveness of 350ms in applications with
25 nodes at 25 FPS, where 350ms represents the time it takes for all nodes to
apply an event since its initial occurrence.
%
As we discuss in Section~\ref{sec.eval}, the performance of our midlleware is
measured in terms of rollbacks, i.e., the more the peers need to resynchronize
and travel back in time, the lower is the performance of the application.
In a similar scenario with 21 peers at 50 FPS, and average of 5 hops with 50ms
latency, we show 3.1\% of rollbacks with responsiveness of 500ms, which we
classify as moderate performance (Figure~\ref{fig.baks}, framed rectangle in
the center).
As expected, the peer-to-peer performance is lower in comparison to centralized
solutions, but still viable in many scenarios, as detalied in
Section~\ref{sec.eval}.
%
Regarding CRDTs, recall that they only provide eventual consistency with no
notion of a shared timeline in which peers go through identical steps.
Therefore, measuring the performance of event responsiveness, maximum FPS, or
clock synchronization is meaningless, given that events can be applied locally
and broadcasted using a best-effort policy.

\subsection{Time Machines}

Regarding decentralized time machines, Fusion~\cite{fusion} is a state
synchronization networking library for the Unity game engine.
It provides networked tick-based simulation in two modes: client-server
(\emph{hosted mode}) and peer-to-peer (\emph{shared mode}).
The hosted mode is based on continuous memory snapshots, which allows for
full rollbacks when clients diverge.
However, the shared mode is less powerful and only supports eventual
consistency among clients.
We presume that continuous memory snapshots would be too costly without a
central server.
%
Regarding single-user local time machines, the video game Braid%
\footnote{Braid: \url{http://braid-game.com/}}
is designed on the assumption that players have unrestricted rollbacks as part
of the game mechanics.
Like Fusion, the implementation is based on continuous memory snapshots,
instead of deterministic simulation as we propose in this work.
%
Schuster~\cite{tml.js} proposes a simple API for time travelling in
event-driven JavaScript:
    \code{init()} to return an initial state;
    \code{handle(evt,mem)} to process events and modify state; and
    \code{render(mem)} to output the current state.
As we describe next, we use a similar approach by separating state mutation
from rendering~\cite{tml.alive}.
